Anony\emph{mice}d shareable data: Using \emph{mice} to create and analyze multiply imputed synthetic data sets

### 2021-10-04 manuscript/Manuscript.tex
MDPI manuscript
cf rticles package

% see https://stackoverflow.com/a/47122900

@@

@@

\hypertarget{version}{%
\section{Version}\label{version}}

This Rmd-skeleton uses the mdpi Latex template published 2019/02.
However, the official template gets more frequently updated than the
`rticles' package. Therefore, please make sure prior to paper
submission, that you're using the most recent .cls, .tex and .bst files
(available \href{http://www.mdpi.com/authors/latex}{here}).

@@

The introduction should briefly place the study in a broad context and
highlight why it is important. It should define the purpose of the work
and its significance. The current state of the research field should be
reviewed carefully and key publications cited. Please highlight
controversial and diverging hypotheses when necessary. Finally, briefly
mention the main aim of the work and highlight the principal
conclusions. As far as possible, please keep the introduction
comprehensible to scientists outside your particular field of research.
Citing a journal paper
\citep{bertrand-krajewski_distribution_1998, leutnant_stormwater_2016}.
And now citing a book reference \citet{gujer_systems_2008}. Please use
the command \citep{leutnant_stormwater_2016} for the following MDPI
journals, which use author-date citation: Administrative Sciences, Arts,
Econometrics, Economies, Genealogy, Humanities, IJFS, JRFM, Laws,
Religions, Risks, Social Sciences.

@@

Materials and Methods should be described with sufficient details to
allow others to replicate and build on published results. Please note
that publication of your manuscript implicates that you must make all
materials, data, computer code, and protocols associated with the
publication available to readers. Please disclose at the submission
stage any restrictions on the availability of materials or information.
New methods and protocols should be described in detail while
well-established methods can be briefly described and appropriately
cited.

Research manuscripts reporting large datasets that are deposited in a
publicly available database should specify where the data have been
deposited and provide the relevant accession numbers. If the accession
numbers have not yet been obtained at the time of submission, please
state that they will be provided during review. They must be provided
prior to publication.

Interventionary studies involving animals or humans, and other studies
require ethical approval must list the authority that provided approval
and the corresponding ethical approval code.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\conflictsofinterest{Declare conflicts of interest or state `The authors
declare no conflict of interest.' Authors must identify and declare any
personal circumstances or interest that may be perceived as
inappropriately influencing the representation or interpretation of
reported research results. Any role of the funding sponsors in the
design of the study; in the collection, analyses or interpretation of
data in the writing of the manuscript, or in the decision to publish the
results must be declared in this section. If there is no role, please
state `The founding sponsors had no role in the design of the study; in
the collection, analyses, or interpretation of data; in the writing of
the manuscript, an in the decision to publish the results'.}

@@

%=====================================
% References, variant B: external bibliography
%=====================================
\reftitle{References}
\externalbibliography{yes}
\bibliography{mybibfile.bib}

>>>>

% see https://stackoverflow.com/a/47122900
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}

@@

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

@@

@@

Open science, including open data, has been marked as the future of
science \citep{gewin_data_2016}, and the advantages of publicly
available research data are numerous
\citep{molloy_open_2011, walport_brest_sharing_2011}. Collecting
research data requires an enormous investment both in terms of time and
monetary resources. Openly accessible research data bears the potential
of increasing the scientific returns for the same data collection
effort. Additionally, the fact that public funds are used for data
collection results in increasing demand for the collected data.
Nevertheless, the possibilities to distribute research data directly are
often very limited due to restrictions on data privacy and data
confidentiality. Although these regulations are much needed, privacy
constraints are also ranked among the toughest challenges to overcome in
the advancement of modern day social science research
\citep{lazer_life_2009}.

Anonymizing research data might seem a quick and appealing approach to
limit the unique identification of participants. However, this approach
is not sufficient to fulfil contemporary privacy and confidentiality
requirements \citep{ohm_broken_2009, national_putting_2007}. Over the
years, several other techniques have been used to increase the
confidentiality of research data, such as categorizing continuous
variables, top coding values above an upper bound or adding random noise
to the observed values \citep{drechsler_synthetic_2011}. However, these
methods may distort the true data relation between variables, thereby
reducing the data quality and the scientific returns for re-using the
same data for further research.

An alternative solution has been proposed separately by
\citet{rubin_statistical_disclosure_1993} and
\citet{little_statistical_1993}. Although their approaches differ to
some extent, the overarching procedure is to use bonafide observed data
to generate multiply imputed synthetic data sets that can be freely
disclosed. While in practice, one could see this as replacing the
observed data values by multiple draws from the posterior predictive
distribution of the observed data, based on some imputation model, Rubin
would argue that these synthetic data values are merely draws from the
same true data generating model. In that sense, the observed data is
never replaced, but the population is resampled from the information
captured in the (incomplete) sample. Using this approach, the researcher
could replace the observed data set as a whole with multiple synthetic
versions. Alternatively, the researcher could opt to only replace a
subset of the observed data. For example, one can choose to only replace
dimensions in the data that could be compared with publicly available
data sets or registers. Likewise, synthetisation could be limited to
those values that are disclosive, such as high incomes or high
turnovers.

Conceptually, the synthetic data framework is based upon the building
blocks of multiple imputation of missing data, as proposed by
\citet{rubin_multiple_1987}. Instead of replacing just the missing
values with multiple draws from the posterior predictive distribution,
one could easily \emph{overimpute} any observed sensitive values.
Similarly to multiple imputation of missing data, the multiple synthetic
data sets allow for correct statistical inferences, despite the fact
that the analyses do not use the ``true'' value. The analyses over
multiple synthetic data sets should be pooled into a single inference,
so that the researcher can draw valid conclusions from the pooled
results. To that respect, the variance should reflect the added
variability that is induced by the imputation procedure.

Potentially, this approach could fulfill the needs for openly accessibly
data, without running into barriers with regard to privacy and
confidentiality constraints. However, there is no such thing as a free
lunch: data collectors have to put effort in creating high-quality
synthetic data. Also, the quality of the synthetic data is highly
dependent on the imputation models, and using flawed models to generate
synthetic data might bias subsequent analyses. Conversely, if the models
used to create the synthetic data are able to preserve the relationships
between the variables as in the original data, the synthetic data can be
nearly as informative as the observed data. Thus, to fully exploit the
benefits of synthetic data, the effort to actually create these
high-quality data sets should be kept at a minimum.

To mitigate additional effort of creating synthetic data sets on behalf
of the researcher, software aimed at multiple imputation of missing data
can be employed. Especially if researchers used this software at an
earlier stage in the research process, or acquired familiarity with it
during earlier projects, the additional burden of creating synthetic
data sets is relatively small. The R-package \texttt{mice} \citep{mice}
implements multiple imputation of missing data in a straightforward and
user-friendly manner. However, the functionality of \texttt{mice} is not
restricted to the imputation of missing data, but allows to impute any
value in the data: even observed values. Consequently, \texttt{mice} can
be utilized for the creation of multiply imputed synthetic data sets.

After creating the multiply imputed synthetic data sets, the goal is to
obtain valid statistical inferences in the spirit of
\citet{rubin_multiple_1987} and \citet{neyman1934}. In the missing data
framework, this is done by performing statistical analyses on all
imputed data sets, and pooling the results of the analyses according to
Rubin's rules \citep[pp.76]{rubin_multiple_1987}. In the synthetic data
framework, the same procedure is followed, but with a slight twist:
there are no values that remain constant over the synthetic data sets.
The procedure of drawing valid inferences from multiple synthetic data
sets is therefore slightly different.

In this manuscript we detail a workflow for synthesizing data with
\texttt{mice}. First, the \texttt{mice} algorithm for the creation of
synthetic data will be shortly explained. The aim is to generate
synthetic sets that reassure the privacy and confidentiality of the
participants. Second, a straightforward workflow for imputation of
synthetic data with \texttt{mice} will be demonstrated. Third, we
demonstrate the validity of the procedure through statistical
simulation.

\hypertarget{generating-synthetic-data-with-mice}{%
\section{\texorpdfstring{Generating synthetic data with
\texttt{mice}}{Generating synthetic data with mice}}\label{generating-synthetic-data-with-mice}}

The \texttt{mice} package \citep{mice} in \texttt{R} \citep{Rproject}
has been developed for multiple imputation of missing data. In this
context, the aim is to replace missing values due to nonresponse by
plausible values from the posterior predictive distribution of the
variable containing the missings. Doing so, \texttt{mice} makes use of
fully conditional specification {[}FCS; \citet{vanbuuren_fully_2006}{]},
which breaks down the multivariate distribution of the data
\(\textbf{Y} = (\textbf{Y}_{obs}, \textbf{Y}_{mis})\) into
\(j = 1, 2, \dots, k\) univariate conditional densities, where \(k\)
denotes the number of columns in the data. Using FCS, a model is
constructed for every incomplete variable and the missing values
\(Y_{j, mis}\) are then imputed with draws from the posterior predictive
distribution of \(P(Y_{j, mis} | \textbf{Y}_{obs}, \theta)\) on a
variable-by-variable basis. Note that the predictor matrix \(Y_{-j}\)
may contain yet imputed values from an earlier imputation step, and thus
will be updated after every iteration. This procedure is applied \(m\)
times, resulting in \(m\) completed data sets
\(\textbf{D} = (\textbf{D}^{(1)}, \textbf{D}^{(2)}, \dots, \textbf{D}^{(m)})\),
with \(\textbf{D}^{(l)} = (\textbf{Y}_{obs}, Y^{(l)}_{mis})\).

In \texttt{mice}, the generation of multiply imputed data sets to solve
for unobserved values is straightforward. The following pseudocode
details the multiple imputation of the \texttt{mice::boys} data set
\citep{fredriks_boys_2000} into the object \texttt{imp} with
\texttt{m\ =\ 10} imputated sets and \texttt{maxit\ =\ 7} iterations for
the algorithm to converge, using the default imputations methods for
each column data class.

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(mice)}
\NormalTok{imp }\OtherTok{\textless{}{-}} \FunctionTok{mice}\NormalTok{(boys, }
            \AttributeTok{m =} \DecValTok{10}\NormalTok{,}
            \AttributeTok{maxit =} \DecValTok{7}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

It is straightforward to extended the imputation approach to generate
synthetic values. Rather than imputing missing data, the observed values
are then replaced by synthetic draws from the posterior predictive
distribution. For simplicity, assume that the data is completely
observed (i.e., \(\textbf{Y} = \textbf{Y}_{obs}\)). Following the
notation of \citet{reiter_raghunathan_multiple_2007}, let for \(n\)
units denote \(Z_i = 1\) if any of the values of unit
\(i = 1, 2, \dots, n\) are to be replaced by imputations, and
\(Z_i = 0\) otherwise, with \(\textbf{Z} = (Z_1, Z_2, \dots, Z_n)\).
Accordingly, the data consists of values that are to be replaced and
values that are to be kept (i.e.,
\(\textbf{Y} = (\textbf{Y}_{rep}, \textbf{Y}_{nrep})\). Now, instead of
imputing \(\textbf{Y}_{mis}\) with draws from the posterior predictive
distribution of \(P(Y_{j, mis} | \textbf{Y}_{obs}, \theta)\) as in the
missing data case, \(\textbf{Y}_{rep}\) is imputed from the posterior
distribution of
\(P(Y^{(l)}_{j, rep} | \textbf{Y}^{(l)}_{-j}, \textbf{Z}, \theta)\),
where \(l\) is an indicator for the synthetic data set
(\(l = 1, 2, \dots, m\)). Note that synthetic values that are imputed at
an earlier step can be used for imputing variable \(j\). This process
results in the synthetic data
\(\textbf{D} = (\textbf{D}^{(1)}, \textbf{D}^{(2)}, \dots, \textbf{D}^{(m)})\).

For example, overimputing synthetic values for both the observed and
missing cells in the \texttt{mice::boys} data set into the object
\texttt{syn}, given the same imputation parameters as before, can be
realized by the following code execution.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{syn }\OtherTok{\textless{}{-}} \FunctionTok{mice}\NormalTok{(boys, }
            \AttributeTok{m =} \DecValTok{10}\NormalTok{,}
            \AttributeTok{maxit =} \DecValTok{7}\NormalTok{, }
            \AttributeTok{where =} \FunctionTok{matrix}\NormalTok{(}\ConstantTok{TRUE}\NormalTok{, }
                           \AttributeTok{nrow =} \FunctionTok{nrow}\NormalTok{(boys),}
                           \AttributeTok{ncol =} \FunctionTok{ncol}\NormalTok{(boys)))}
\end{Highlighting}
\end{Shaded}

where the argument \texttt{where} requires a matrix of the same
dimensions as the data, (i.e., a \(n \times k\) matrix) containing
logicals \(z_{ij}\) that indicate which cells are selected to have their
values replaced by draws from the posterior predictive distribution.
This approach allows to \emph{overimpute} a subset of the observed data,
or - as in the above example - the observed data as a whole, resulting
in a data set that partly or completely consists of synthetic data
values.

Choosing an adequate imputation model to impute the data is paramount,
as a flawed imputation model may drastically impact the validity of
inferences. Imputation models should be as flexible as possible to
capture most of the patterns in the data, and to model possibly
unanticipated data characteristics
\citep{murray_multiple_2018, rubin_18years_1996}. Parametric methods,
albeit easy to implement in practice, may be too restrictive to capture
generally complex patterns in the data, especially in the case of
nonlinear relations and interactions between multiple variables.
Classification and regression trees {[}CART;
\citet{breiman_cart_1984}{]} allow to model more complex patterns in the
data, and have therefore been suggested as an appropriate imputation
method
\citep{reiter_cart_2005, burgette_reiter_cart_2010, doove_buuren_recursive_2014}.
Loosely speaking, CART sequentially splits the predictor space into
non-overlapping regions in such a way that the within-region variance is
as small as possible after every split. As such, CART does not impose
any parametric distribution on the data, making it a widely applicable
method that allows for a large variety of relationships within the data
\citep{islr_2013}. Given these appealing characteristics and the call
for the use of flexible methods when multiply imputing data, we will
focus our illustrations and evaluations of \texttt{mice} to method
\texttt{mice.impute.cart()}, realized by:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{syn }\OtherTok{\textless{}{-}} \FunctionTok{mice}\NormalTok{(boys, }
            \AttributeTok{m =} \DecValTok{10}\NormalTok{,}
            \AttributeTok{maxit =} \DecValTok{7}\NormalTok{, }
            \AttributeTok{method =} \StringTok{"cart"}\NormalTok{,}
            \AttributeTok{where =} \FunctionTok{matrix}\NormalTok{(}\ConstantTok{TRUE}\NormalTok{, }
                           \AttributeTok{nrow =} \FunctionTok{nrow}\NormalTok{(boys),}
                           \AttributeTok{ncol =} \FunctionTok{ncol}\NormalTok{(boys)))}
\end{Highlighting}
\end{Shaded}

In a nutshell, the above code shows the simplicity of creating
\(m = 10\) synthetic data sets using \texttt{mice}. In practice,
however, one should take some additional complicating factors into
account. For example, one should account for deterministic relations in
the data. Additionally, relations between variables may be described
best using a different model than \texttt{CART}. Such factors are data
dependent and should be considered by the imputer. In the next section,
we will describe how the \texttt{boys} data can be adequately imputed.
Additionally, we will show through simulations that this approach yields
valid inferences.

@@

We demonstrate the suitability of using \texttt{mice} for synthesization
using a simulation study on the \texttt{mice::boys} data set. This data
set consists of the values of \(748\) Dutch boys on the following \(9\)
variables:

\begin{longtable}[]{@{}ll@{}}
\toprule
column & description \\
\midrule
\endhead
age & age in years \\
hgt & height (cm) \\
wgt & weight (kg) \\
bmi & body mass index \\
hc & head circumference (cm) \\
gen & genital Tanner stage G1-G5 \\
phb & pubic hair Tanner P1-P6 \\
tv & testicular volume (ml) \\
reg & region \\
\bottomrule
\end{longtable}

Unfortunately, this data set does not differ from the vast majority of
collected data sets, in the sense that it suffers from missing data. For
simplicity, the data is completed using the default \texttt{mice}
imputation model for all predictors except \texttt{bmi}, which is
passively imputed using its deterministic relation with weight and
height.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# create a single imputed, completely observed \textasciigrave{}boys\textasciigrave{} data set}
\FunctionTok{set.seed}\NormalTok{(}\DecValTok{123}\NormalTok{)}

\NormalTok{meth }\OtherTok{\textless{}{-}} \FunctionTok{make.method}\NormalTok{(boys)}
\NormalTok{meth[}\StringTok{"bmi"}\NormalTok{] }\OtherTok{\textless{}{-}} \StringTok{"\textasciitilde{} I(wgt / (hgt / 100)\^{}2)"}
\NormalTok{pred }\OtherTok{\textless{}{-}} \FunctionTok{make.predictorMatrix}\NormalTok{(boys)}
\NormalTok{pred[}\FunctionTok{c}\NormalTok{(}\StringTok{"hgt"}\NormalTok{, }\StringTok{"wgt"}\NormalTok{), }\StringTok{"bmi"}\NormalTok{] }\OtherTok{\textless{}{-}} \DecValTok{0}

\NormalTok{imp }\OtherTok{\textless{}{-}} \FunctionTok{mice}\NormalTok{(boys, }
            \AttributeTok{m =} \DecValTok{1}\NormalTok{,}
            \AttributeTok{maxit =} \DecValTok{10}\NormalTok{,}
            \AttributeTok{method =}\NormalTok{ meth,}
            \AttributeTok{predictorMatrix =}\NormalTok{ pred)}

\NormalTok{data }\OtherTok{\textless{}{-}} \FunctionTok{complete}\NormalTok{(imp)}
\end{Highlighting}
\end{Shaded}

\hypertarget{simulation-methods}{%
\subsection{Simulation methods}\label{simulation-methods}}

To induce sampling variance, 1000 bootstrap samples of the \texttt{boys}
data will be synthesized with \(m = 5\) imputations. Synthetic values
are generated using the \texttt{CART} imputation method for all columns,
except for \texttt{bmi}. The deterministic relation \texttt{bmi} which
will be synthesized passively based on the synthetic values for
\texttt{hgt} and \texttt{wgt} to preserve the relation in the synthetic
data. Additional parameters that come with the use of
\texttt{mice.impute.cart()} are the complexity parameter \texttt{cp} and
the minimum number of observations in any terminal node
\texttt{minbucket}, that both constrain the flexibility of the
imputation model. The values of the parameters \texttt{cp} and
\texttt{minbucket} ought to adhere to the call for imputation models
that are as flexible as possible. Appropriate values for these
parameters, as well as the input for the \texttt{predictorMatrix},
depend on the data at hand. In the current example, the complexity
parameter is specified at \texttt{cp\ =\ 1e-08} rather than the default
value \texttt{1e-04}, and the minimum number of observations in each
terminal node is set at \texttt{minbucket} \(= 3\) rather than the
default value \(5\). By allowing for more complexity in the imputation
model, bias in the estimates from the synthetic data set is reduced.
Additionally, since the missingness pattern is monotone, the number of
iterations can be set to \texttt{maxit\ =\ 1}.

To assess the performance of \texttt{mice} for synthesizing data, we
compare the bootstrapped samples with the synthetic versions of these
bootstrapped samples. Specifically, univariate descriptive statistics,
the correlation matrix, and two linear regression models as well as one
ordered logistic regression model will be considered. Subsequently, the
bias in the parameters and the \(95\%\) confidence interval coverage of
the synthetic data will be examined. Similarly to multiple imputation of
missing data, correct inferences from synthetic data requires correct
pooling over the multiply imputed data sets.

Obtaining a final point estimate of the parameter of interest \(Q\)
after imputation is fairly easy and no different from pooling in the
case of missing data \citep{rubin_multiple_1987}. One can calculate the
average of the \(m\) point estimates \(q^{(l)}\) \[
\bar{q}_m = \sum_{l = 1}^m \frac{q^{(l)}}{m}.
\] with \(l = 1, \dots, m\).

Similarly to the missing data case, variances, and subsequently
confidence intervals, should incorporate the increase in variance that
is due to imputation
\citep{reiter_partially_inference_2003, drechsler_synthetic_2011}. Yet,
the increase in variance due to imputation differs according to whether
missing values are imputed or observed data is overimputed with missing
values. Whereas the variance estimate after imputation of missing data
needs to account for the fact that a certain amount of information in
the data is missing, variance estimation from synthetic data does not
suffer from this issue. The adjusted variance estimate that follows from
using multiple synthetic data sets only suffers from the fact that a
finite number of \(m\) synthetic data sets are used to resemble the
observed data. Hence, the according variance estimate for synthetic data
as developed by \citet{reiter_partially_inference_2003} yields \[
T = \bar{u}_m + \frac{b_m}{m},
\]

with between-imputation variance

\[
b_m = \sum_{l = 1}^m \frac{(q^{(l)} - \bar{q}_m)^2}{(m - 1)}, \\
\] and sampling variance

\[
\bar{u}_m = \sum_{l = 1} \frac{u^{(l)}}{m},
\] with \(u^{(l)}\) reflecting the variance estimate in the \(l\)th
synthetic data set.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\conflictsofinterest{`The authors declare no conflict of interest.'}

@@

%=====================================
% References, variant B: external bibliography
%=====================================
\reftitle{References}
\externalbibliography{yes}
\bibliography{synth.bib}
